Road Segment Prioritization for Bicycle Infrastructure

### 2021-01-04 README.Rmd

---
title: Road Segment Prioritization for Bicycle Infrastructure
date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport. The aim of this paper is to develop methods to support the prioritisation of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally. Distance decay, routing, and network analysis techniques are used to examine where future cycling demand is most likely to arise, how such demand could be accomodated within existing street networks, and how to ensure a fair distribution of investment. The city of Manchester is used as a case study to demonstrate the methods which, as demonstrated in reproducible code and open access data accompanying the paper, can be applied to any city. We find that flexible routing profiles that are adjustable to local needs can avoid high-stress road segments, enabling the identification of networks that balance directness against the need for safe and stress-free routes. Community detection is used to partition the city and evaluate results from 'utilitarian' and 'egalitarian' policy perspectives. Based on discussion of the findings from a policy perspective, we conclude that the 'egalitarian' model prioritisation approach demonstrated in this paper has great potential to support cost-effective strategic cycle network planning internationally. \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
output:
  # bookdown::github_document2:
  # bookdown::pdf_document2:
  bookdown::word_document2:
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
---

@@

The analysis is heavily dependent on commuter data^[*commuter data* refers to publicly available Origin-Destination (OD) data, normally obtained from a census. This study uses data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between MSOA pairs, by mode of travel.]. Commuter data in the UK is publicly available at the Middle layer Super Output Area (MSOA) level; the average MSOA has a population of 8209 [@ofn2018population]. @iacono2010measuring note that such large travel zones are not ideal for understanding route choice behaviour of cyclists and pedestrians. They also give rise to an \`ecological fallacy' whereby average characteristics are assumed to apply to all residents of the aggregated geographical area. Given that more granular data is not publicly available, the study uses MSOA-level commuter data. The methodology is however applicable to more granular commuter data should it become available.

@@

# Conclusions

<!-- I think this section should start with a short (one sentence) re-statement of what the paper has done. -->
<!-- (HM) Answer: done --> 
This paper provides a methodology for prioritizing road segments for investments in cycling infrastructure, with the output being a proposed cycling network for the city under study.

The methodology is designed so as to respect both the needs of the users and the ambitions of transport planners and local/regional level government stakeholders. It can can be used as a starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data. The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the national census)^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data. Instructions for reproducing the results are detailed in this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)]. It should be noted that the recommendations are only as good as the granularity of the available data.

>>>>

---
title: Road Segment Prioritization for Bicycle Infrastructure
date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport. The aim of this paper is to develop methods to support the prioritisation of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally. Distance decay, routing, and network analysis techniques are used to examine where future cycling demand is most likely to arise, how such demand could be accomodated within existing street networks, and how to ensure a fair distribution of investment. The city of Manchester is used as a case study to demonstrate the methods which, as demonstrated in reproducible code and open access data accompanying the paper, can be applied to any city. We find that flexible routing profiles that are adjustable to local needs can avoid high-stress road segments, enabling the identification of networks that balance directness against the need for safe and stress-free routes. Community detection is used to partition the city and evaluate results from 'utilitarian' and 'egalitarian' policy perspectives. Based on discussion of the findings from a policy perspective, we conclude that the 'egalitarian' model prioritisation approach demonstrated in this paper has great potential to support cost-effective strategic cycle network planning internationally. \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
output:
  # bookdown::github_document2:
  bookdown::pdf_document2:
  # bookdown::word_document2:
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
---

@@

The approach is based on origin-destination (OD), which can be obtained from many sources.
In this paper we use open access data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between Middle layer Super Output Area (MSOA) zones, which have an average just over 8000 [@ofn2018population].

@@

# Disussion and conclusions

This paper provides a methodology for prioritizing road segments for investments in cycling infrastructure, with the output being a proposed cycling network for the city under study. 
The approach aims to respect both the needs of the users and the ambitions of stakeholders working at local/regional levels.
The results, primarily detailed route network maps based on current travel behaviour derived from origin-destination data, can starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data. The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the national census)^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data. Instructions for reproducing the results are detailed in this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)].

The approach is not without limitations.
The level of detail is only as good as the granularity of the available data (in this case relatively coarse zones).
@iacono2010measuring note that such large travel zones are not ideal for understanding route choice behaviour of cyclists and pedestrians. They also give rise to an \`ecological fallacy' whereby average characteristics are assumed to apply to all residents of the aggregated geographical area, suggesting a need to applying the methods to more granular origin-destination data (and for governments and other data-collecting organisations to make origin-destination more readily available).
